### 2011-12-07 report/paper.tex
Renamed paper report

% Matthew Monaco
% Andy Sayler
% Landon Spear
% University of Colorado
% Networked Character devices
% Fall 2011

\documentclass[11pt,twocolumn]{article}

\usepackage[text={6in, 9in}, centering]{geometry}
\usepackage{graphicx}
\usepackage{url}

\bibliographystyle{plain}

\newenvironment{packed_enum}{
\begin{enumerate}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{enumerate}}

\newenvironment{packed_item}{
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

\begin{document}

\title{Networked Character Devices}
\author{Monaco, Matthew \and Sayler, Andrew \and Spear, Landon
  \\ \and University of Colorado}
\date{\today}

\maketitle

\begin{abstract}
Lorem ipsum
\end{abstract}

% Landon Sections

\section{Introduction}
\label{sec:introduction}

We did some stuff, we used some references \cite{ldd3}.

\section{Related Work}
\label{sec:relatedwork}

We are unique.

% Matt Sections

% Combine Architecture and System?
\section{Architecture}
\label{sec:architecture}

\section{System}
\label{sec:system}

\section{Implementation}
\label{sec:implementation}

% Andy Sections

\section{Results and Evaluation}
\label{sec:results}

The Network Character Device system is still a work in
progress. We will need to Expand the system beyond it's current state
before the NCD concept is ready for production-level use.

Non the less, we have gained many insights into the advantages such a
system might provide. We have also encountered the difficulties
implementing such a system into an existing OS like Linux entails.

\subsection{Current Status}
\label{sec:currentstatus}

As of this writing, we have completed a partial implementation of
the Network Character Device system. Current functionality includes:

\begin{packed_item}
\item Support for exporting a single character device from the server
\item Support for importing a single character device on the client
\item Client-side kernel-space Linux module implementation
\item Server-side user-space background process implementation
\item Support for open, close, read, and write calls
\item Basic Linux udev support for automatic node creation on client
\end{packed_item}

This set of core functionality provides a working proof-of-concept
Network Character Device system. This system fully supports the
export of a single character device from server to client over the
network and demonstrates the basic use and utility of such a system.

It should be noted, however, that some work is still required before
the Network Character device system can be considered a production
ready utility. Most notably, the following features are necessary for
a fully functional NCD system, but are not yet included in our
implementation:

\begin{packed_item}
\item Multi-device, multi-server client-side import support
\item Multi-device server-side export support
\item Support for exclusive use and protection of exported devices
\item Support for ioctl calls
\item Support for providing and obtaining metadata for exported
  devices
\item Support for ``advertisement'' of available exported devices
\item Support for integrating imported devices into other kernel
  subsystems (human interface subsystem, audio subsystem, etc)
\item Addition of ``NCD-admin'' utilities for managing and administering
  NCD system
\end{packed_item}

Section \ref{sec:futurework} provides some insight into the addition
of some of these expanded features.

\subsection{Advantages}
\label{sec:advantages}

\subsection{Challenges}
\label{sec:challenges}

\section{Future Work}
\label{sec:futurework}

\section{Conclusion}
\label{sec:conclusion}

\bibliography{refs}

\end{document}

### 2011-12-07 report/paper.tex
Added LaTeX Listings support and sample to paper and presentation

@@
 \usepackage{url}
+\usepackage{listings}
 
 \bibliographystyle{plain}
+
+\lstloadlanguages{C}
+\lstset{
+  language=C,
+  basicstyle=\footnotesize,
+  numbers=none,
+  numberstyle=\footnotesize,
+  stepnumber=1,
+  numbersep=5pt,
+  showspaces=false,
+  showstringspaces=false,
+  showtabs=false,
+  tabsize=4,
+  captionpos=b,
+  breaklines=true,
+  breakatwhitespace=false,
+  title=\lstname,
+  frame=single,
+  frameround=tttt
+}
 
@@
 \label{sec:architecture}
+
+\lstinputlisting[float=*,
+  caption={Listing Code Sample},
+  label=lst:codeSample.c]{codesample.c}
 

### 2011-12-07 report/presentation.tex
Added LaTeX Listings support and sample to paper and presentation

% presentation.tex
% Matthew Monaco
% Andy Sayler
% Landon Spear

\documentclass[xcolor=dvipsnames]{beamer}
%\useoutertheme{default}
%\usetheme{Boadilla}
\setbeamercovered{transparent=25}
\setbeamertemplate{blocks}[rounded][shadow=false] 
\setbeamertemplate{navigation symbols}{}

\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}

\bibliographystyle{plain}

\lstloadlanguages{C}
\lstset{
  language=C,
  basicstyle=\footnotesize,
  numbers=none,
  numberstyle=\footnotesize,
  stepnumber=1,
  numbersep=5pt,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=4,
  captionpos=b,
  breaklines=true,
  breakatwhitespace=false,
  title=\lstname,
  frame=single,
  frameround=tttt
}

\title[NCD]{Networked Character Devices}
%\subtitle[]{}
\author[ M. Monaco \& A. Sayler \& L Spear]{Matthew Monaco \&
                                            Andy Sayler \&
                                            Landon Spear}
\institute[University of Colorado]{
  University of Colorado\\
  \texttt{matthew.monaco@colorado.edu}\\*
  \texttt{andrew.sayler@colorado.edu}\\*
  \texttt{landon.spear@colorado.edu}
}
\date[Dec. 10, 2011]{Saturday, December 10\textsuperscript{th}, 2011}

\begin{document}

%---Title Slide---%
\begin{frame}[plain]
  \titlepage
\end{frame}

%Landon - Beginning

\section{Overview}
%---Intro Slide---%
\begin{frame}{\bf NCD Overview}
  We did some stuff, we used some references \cite{ldd3}.
\end{frame}

% Matt - Middle

%---Code Sample Slide---%
\begin{frame}{\bf Code Sample}

\lstinputlisting[
  caption={Listing Code Sample},
  label=lst:codeSample.c]{codesample.c}

\end{frame}

% Andy - End

\section{Bibliography}
%---Bibliography---%
\begin{frame}{\bf Bibliography}

\bibliography{refs}

\end{frame}

\end{document}

### 2011-12-07 report/codesample.c

@@ -1,0 +1,10 @@
+/* This is a Listings test */
+int main(){
+
+	int i = 0;
+	for(i=0; i<10; i++){
+		printf("This is a LaTeX test");
+	}
+
+	return 0;
+}

### 2011-12-07 report/presentation.tex
Added to presentation

@@
 
+\newenvironment{packed_enum}{
+\begin{enumerate}
+  \setlength{\itemsep}{1pt}
+  \setlength{\parskip}{0pt}
+  \setlength{\parsep}{0pt}
+}{\end{enumerate}}
+
+\newenvironment{packed_item}{
+\begin{itemize}
+  \setlength{\itemsep}{1pt}
+  \setlength{\parskip}{0pt}
+  \setlength{\parsep}{0pt}
+}{\end{itemize}}
+
 \title[NCD]{Networked Character Devices}
@@
 
+\section{Introduction}
+
+\section{Related Work}
+
 % Matt - Middle
+
+% Combine Architecture and System?
+\section{Architecture}
+
+\section{System}
+
+\section{Implementation}
 
@@
 
+% Andy Sections
+
+\section{Results and Evaluation}
+%---Current State Slide---%
+\begin{frame}{\bf Current State}
+
+\begin{packed_item}
+\item Supports exporting a single character device from the server
+\item Supports importing a single character device on the client
+\item Supports open, close, read, and write calls
+\item Supports basic Linux udev operation for automatic node creation on client
+\end{packed_item}
+
+\end{frame}
+
+%---ToDo Slide---%
+\begin{frame}{\bf ToDo}
+
+\begin{packed_item}
+\item Multi-device, multi-server client-side import support
+\item Multi-device server-side export support
+\item Support for exclusive use and protection of exported devices
+\item Support for ioctl calls
+\item Support for providing and obtaining metadata for exported
+  devices
+\item Support for ``advertisement'' of available exported devices
+\item Support for integrating imported devices into other kernel
+  subsystems (human interface subsystem, audio subsystem, etc)
+\item Addition of ``NCD-admin'' utilities for managing and administering
+  NCD system
+\end{packed_item}
+
+\end{frame}
+
+%---Advantages Slide---%
+\begin{frame}{\bf Advantages}
+
+\end{frame}
+
+%---Challenges Slide---%
+\begin{frame}{\bf Challenges}
+
+\end{frame}
+
+\section{Future Work}
+
+\section{Conclusion}
+
 \section{Bibliography}

### 2011-12-08 report/presentation.tex
presentation: setup themes, cleanup
presentation: high level arch

@@
 \documentclass[xcolor=dvipsnames]{beamer}
-%\useoutertheme{default}
-%\usetheme{Boadilla}
+\usetheme{AnnArbor}
+\usecolortheme{beaver}
 \setbeamercovered{transparent=25}
@@
 %\subtitle[]{}
-\author[ M. Monaco \& A. Sayler \& L Spear]{Matthew Monaco \&
+\author[Monaco, Sayler, Spear]{Matthew Monaco \&
                                Andy Sayler \&
                                Landon Spear}
-\institute[University of Colorado]{
-  University of Colorado\\
-  \texttt{matthew.monaco@colorado.edu}\\*
-  \texttt{andrew.sayler@colorado.edu}\\*
-  \texttt{landon.spear@colorado.edu}
+\institute[CU-Boulder]{
+  University of Colorado at Boulder \\
+  \texttt{\{first.last\}@colorado.edu}
 }
@@
   \titlepage
+\end{frame}
+
+\begin{frame}{Outline}
+  \tableofcontents
 \end{frame}
@@
 
-\section{System}
+\begin{frame}[c]{Very High Level}
+  \includegraphics[width=0.8\textwidth]{arch-01.png}
+
+  Our system uses a basic client server model
+\end{frame}
 
@@
 
-% Andy - End
+% Matt - End
 
@@
 \end{document}
+
+% vim: set sw=2 ts=2 et : %

### 2011-12-08 report/presentation.tex
presentation: high level arch
presentation: server medium level

@@
 
-  Our system uses a basic client server model
+  \begin{itemize}
+    \item<1-> Our system uses a basic client server model
+    \item<2-> The client initiates all exchanges, the server replies
+  \end{itemize}
+\end{frame}
+
+\begin{frame}[c]{High Level}
+  \includegraphics[width=0.8\textwidth]{arch-02.png}
+
+  \begin{itemize}
+    \item<1-> The server side resides entirely in a single kernel module
+    \item<2-> The client side is a standard userspace server
+  \end{itemize}
+\end{frame}
+
+\begin{frame}[c]{Server Medium Level}
+  \includegraphics[width=0.4\textwidth]{arch-03.png}
+
+  \begin{itemize}
+    \item<1-> By default, devices are exposed under /dev/netchar
+    \item<2-> In the future, we should be able to choose any name
+  \end{itemize}
 \end{frame}

### 2011-12-08 report/presentation.tex
presentation: typo s/Server/Client/
Added system images

@@
 \author[Monaco, Sayler, Spear]{Matthew Monaco \&
-                               Andy Sayler \&
+                               Andrew Sayler \&
                                Landon Spear}
@@
 \section{Results and Evaluation}
+%---Full State Slide---%
+
+\begin{frame}[c]{Full System}
+  \begin{center}
+    \includegraphics[height=0.75\paperheight,keepaspectratio]{system-full.png}
+  \end{center}
+\end{frame}
+
+%---Current State Image Slide---%
+\begin{frame}[c]{Current State}
+  \begin{center}
+    \includegraphics[height=0.75\paperheight,keepaspectratio]{system-working.png}
+  \end{center}
+\end{frame}
+
 %---Current State Slide---%
@@
 
-\begin{packed_item}
-\item Supports exporting a single character device from the server
-\item Supports importing a single character device on the client
-\item Supports open, close, read, and write calls
-\item Supports basic Linux udev operation for automatic node creation on client
-\end{packed_item}
+\begin{itemize}
+\item<1-> Supports exporting a single character device from the server
+\item<2-> Supports importing a single character device on the client
+\item<3-> Supports open, close, read, and write calls
+\item<4-> Supports basic Linux udev operation for automatic node creation on client
+\end{itemize}
 
@@
 
-\begin{packed_item}
+\begin{itemize}
 \item Multi-device, multi-server client-side import support
@@
   NCD system
-\end{packed_item}
+\end{itemize}
 

### 2011-12-09 report/presentation.tex
presentation: server medium level
Added a few slides...

--- 2011-12-08 report/presentation.tex
@@
 
-\begin{frame}[c]{Server Medium Level}
+\begin{frame}[c]{Client Medium Level}
   \includegraphics[width=0.4\textwidth]{arch-03.png}
@@
     \item<2-> In the future, we should be able to choose any name
+  \end{itemize}
+\end{frame}
+
+\begin{frame}[c]{Server Medium Level}
+  \includegraphics[width=0.8\textwidth]{arch-02.png}
+  \begin{itemize}
+    \item<1-> Listen on a specified port
+    \item<2-> Receive requests from client
+    \begin{itemize}
+      \item<2-> open,close,read,write,...
+    \end{itemize}
+    \item<3-> Can actually export \textit{any} file
+    \item<4-> Crashes ungracefully when the client disconnects
   \end{itemize}
@@
 
-\begin{packed_item}
+\begin{itemize}
 \item Supports exporting a single character device from the server
@@
 \item Supports open, close, read, and write calls
-\item Supports basic Linux udev operation for automatic node creation on client
-\end{packed_item}
+\item Supports basic Linux udev operation for automatic node
+  creation on client
+\item Tested with various ``real'' and device files
+\end{itemize}
 
@@
   NCD system
-\end{packed_item}
+\item Addition of graceful connect and disconnect handling in real-time
+\end{itemize}
 
@@
 
+What are the benefits of using Networked Character Device?
+
+\begin{itemize}
+\item Divorces peripherals from the computers to which they are
+  attached
+\item Allows simple creation of a ``Device Server''
+\item Extends the ``Everything Is A File'' -nix Philosophy
+\item Allows more efficient use of devices
+\item Allows devices to be networked in a generalized, uniform manner
+\end{itemize}
+
 \end{frame}
@@
 
+But there are difficulties making it work on a production scale...
+
+\begin{itemize}
+\item Does not work well when Character Device is not the exclusive
+  device interface
+  \begin{itemize}
+  \item Many modern day devices interface directly with other
+    in-kernel function, interfaces, and APIs
+  \item procfs and sysfs provide additional device interface points
+  \item Socket interfaces (netlink, etc) provide alternative to file
+    system interfaces
+  \end{itemize}
+\item Requires exported devices to be ``de-integrated'' from host system
+\item Requires QOS guarantees and high guilty network connection
+\end{itemize}
+
+\end{frame}
+
+%---Potential Uses Slide---%
+\begin{frame}{\bf Potential Uses}
+
+What can we use NCDs for?
+
+\begin{itemize}
+\item Quick, cheap, and simple ad-hoc network KVM systems
+\item Remote, networked web-cams, scanners, printers, etc
+\item Basis of ``protocol''-over-Ethernet systems (USB-over-Ethernet,
+  SATA-over-Ethernet, Etc)
+\item Basis of redundant, ``cloud-like'' device pools
+\item Basis of peripheral cluster systems (GPU Clusters, RNG Clusters,
+  Etc)
+\end{itemize}
+
 \end{frame}
@@
 \section{Future Work}
+%---Potential Uses Slide---%
+\begin{frame}{\bf Future Work}
+
+\begin{itemize}
+\item Add multi-device, multi-server support
+\item Add kernel sub-system handler layer (HID, etc)
+\item Add device protection and removal capabilities to server system
+\item Add advertisement protocol
+\item Consider removing networking code from client kernel module
+\end{itemize}
+
+\end{frame}
 

### 2011-12-09 report/presentation.tex
presentation: center images
presentation: remove code sample
presentation: add protocol architecture frame
presentation: use fixed font for some terms, vimspell
presentation: Most of Protocol and Client subsections

@@
 \begin{frame}[c]{Very High Level}
+  \begin{center}
     \includegraphics[width=0.8\textwidth]{arch-01.png}
+  \end{center}
 
@@
 \begin{frame}[c]{High Level}
+  \begin{center}
     \includegraphics[width=0.8\textwidth]{arch-02.png}
+  \end{center}
 
@@
 \begin{frame}[c]{Client Medium Level}
+  \begin{center}
     \includegraphics[width=0.4\textwidth]{arch-03.png}
+  \end{center}
 
   \begin{itemize}
-    \item<1-> By default, devices are exposed under /dev/netchar
+    \item<1-> By default, devices are exposed under \texttt{/dev/netchar}
     \item<2-> In the future, we should be able to choose any name
@@
 \begin{frame}[c]{Server Medium Level}
+  \begin{center}
     \includegraphics[width=0.8\textwidth]{arch-02.png}
+  \end{center}
   \begin{itemize}
@@
     \begin{itemize}
-      \item<2-> open,close,read,write,...
+      \item<2-> \texttt{open()},\texttt{close()},\texttt{read()},\texttt{write()},...
     \end{itemize}
@@
 
+\begin{frame}[c]{Protocol}
+  \begin{center}
+    \includegraphics[width=0.8\textwidth]{proto.png}
+  \end{center}
+
+  \begin{itemize}
+    \item<1-> Structures passed through \texttt{INET} socket
+    \item<2-> Raw data immediately follows when necessary
+    \begin{itemize}
+      \item<2-> \texttt{read()}, \texttt{write()}, ...
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
 \section{Implementation}
 
-%---Code Sample Slide---%
-\begin{frame}{\bf Code Sample}
+\section{Protocol}
 
-\lstinputlisting[
-  caption={Listing Code Sample},
-  label=lst:codeSample.c]{codesample.c}
+\begin{frame}[fragile]{Protocol}
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+      \begin{lstlisting}[gobble=8]
+        enum FOPS {
+          FOP_OPEN, FOP_RELEASE,
+          FOP_READ, FOP_WRITE,
+          _FOP_ERROR
+        };
 
+        struct fop_request {
+
+          enum FOPS call;
+
+          union {
+            struct {
+              int     flags;
+              mode_t  mode;
+            };
+
+            size_t  count;
+          };
+        };
+      \end{lstlisting}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \begin{lstlisting}[gobble=8]
+        struct fop_reply {
+
+          enum FOPS call;
+
+          union {
+            int      open;
+            int      close;
+            ssize_t  read;
+            ssize_t  write;
+          };
+        };
+      \end{lstlisting}
+    \end{column}
+  \end{columns}
 \end{frame}
+
+\subsection{Client}
+
+\begin{frame}[fragile]{Hello, World!}
+  \begin{columns}
+    \begin{column}{0.6\textwidth}
+      \begin{lstlisting}[gobble=8]
+        #include <linux/kernel.h>
+        #include <linux/module.h>
+
+        static int __init nc_init(void)
+        {
+          /* do stuff */
+          return 0;
+        }
+
+        static int __exit nc_exit(void)
+        {
+          /* undo stuff */
+        }
+
+        MODULE_LICENSE("GPL");
+
+        module_init(nc_init);
+        module_exit(nc_exit);
+
+      \end{lstlisting}
+    \end{column}
+    \begin{column}{0.4\textwidth}
+      \begin{itemize}
+        \item initial setup
+        \item cleanup important!
+      \end{itemize}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}[fragile]{Character Device Driver}
+  \begin{lstlisting}[gobble=4]
+    #include <linux/cdev.h>
+    #include <linux/device.h>
+    #include <linux/fs.h>
+
+    static dev_t          nc_dev_t;
+    static struct class*  nc_class;
+    static struct cdev*   nc_cdev;
+    static struct device* nc_device;
+
+  \end{lstlisting}
+
+  \begin{itemize}
+    \item<1-> \texttt{dev\_t}: major/minor number
+    \item<2-> \texttt{class}: subsystem
+    \item<3-> \texttt{cdev}: handler for character devices
+    \item<4-> \texttt{device}: userspace \texttt{/dev} node
+    \item<5-> each has some form of registration mechanism
+  \end{itemize}
+\end{frame}
+
+\begin{frame}[fragile]{Filesystem Handler}
+  \begin{lstlisting}[gobble=4]
+    #include <linux/fs.h>
+
+    static int netchar_open(struct inode* inp, struct file* fp);
+    static ssize_t netchar_read(struct file* fp, char* buffer,
+                                size_t length, loff_t* offset);
+
+    static struct file_operations nc_fops = {
+      /* ... */
+      .open  = nc_open,
+      .read  = nc_read,
+    }
+  \end{lstlisting}
+
+  \begin{itemize}
+    \item<1-> required when registering character device
+    \item<2-> buffer is a userspace pointer!
+    \item<3-> event driven, blocking, asynchronous
+  \end{itemize}
+\end{frame}
+
+\begin{frame}[fragile]{Socket Client}
+  \begin{itemize}
+    \item<1-> socket type: \texttt{AF\_INET}, \texttt{SOCK\_STREAM}, \texttt{IPPROTO\_TCP}
+    \begin{itemize}
+      \item<2-> should be \texttt{AF\_NETLINK}
+      \item<2-> do not want to lose messages
+    \end{itemize}
+    \item<3-> no familiar \texttt{read()}, \texttt{write()}
+  \end{itemize}
+\end{frame}
+
+\subsection{Server}
+
+\subsection{Misc}
 
@@
 
-% vim: set sw=2 ts=2 et : %
+% vim: set sw=2 ts=2 et spell : %

### 2011-12-09 report/presentation.tex
presentation: protocol is a subsection, vim sts
presentation: server subsection and stubs for misc

@@
 
-\section{Protocol}
+\subsection{Protocol}
 
@@
 
-\begin{frame}[fragile]{Socket Client}
+\begin{frame}{Socket Client}
   \begin{itemize}
@@
 
+\begin{frame}[fragile]{Server (\texttt{read()})}
+  \begin{lstlisting}[gobble=4]
+    struct fop_request req;
+    struct fop_reply   rep;
+    while (1) {
+      err = read(sd, &req, sizeof(req));
+      switch(req.call) {
+      case FOP_READ:
+        payload = realloc(payload, req.count);
+        rep.read = read(fd, payload, req.count);
+        if (rep.read < 0)
+          rep.read = -errno;
+        break;
+      }
+      err = write(sd, &rep, sizeof(rep));
+      if (rep.call == FOP_READ && rep.read > 0)
+        err = write(sd, payload, rep.read)
+    }
+  \end{lstlisting}
+\end{frame}
+
 \subsection{Misc}
+
+\begin{frame}{\texttt{udev}}
+\end{frame}
+
+\begin{frame}{\texttt{ncadmin}}
+\end{frame}
 
@@
 
-% vim: set sw=2 ts=2 et spell : %
+% vim: set sw=2 ts=2 sts=2 et spell : %

### 2011-12-09 report/presentation.tex
Added a few images and slides, updated ref with full auther names

@@
 
-% Combine Architecture and System?
 \section{Architecture}
@@
   \begin{itemize}
-\item Supports exporting a single character device from the server
-\item Supports importing a single character device on the client
-\item Supports open, close, read, and write calls
-\item Supports basic Linux udev operation for automatic node
+  \item<1-> Supports exporting a single character device from the server
+  \item<2-> Supports importing a single character device on the client
+  \item<3-> Supports open, close, read, and write calls
+  \item<4-> Supports basic Linux udev operation for automatic node
     creation on client
-\item Tested with various ``real'' and device files
+  \item<5-> Tested with various ``real'' and device files
   \end{itemize}
@@
 \begin{itemize}
-\item Multi-device, multi-server client-side import support
-\item Multi-device server-side export support
-\item Support for exclusive use and protection of exported devices
-\item Support for ioctl calls
-\item Support for providing and obtaining metadata for exported
+\item<1-> Multi-device, multi-server client-side import support
+\item<2-> Multi-device server-side export support
+\item<3-> Support for exclusive use and protection of exported devices
+\item<4-> Support for ioctl calls
+\item<5-> Support for providing and obtaining metadata for exported
   devices
-\item Support for ``advertisement'' of available exported devices
-\item Support for integrating imported devices into other kernel
+\item<6-> Support for ``advertisement'' of available exported devices
+\item<7-> Support for integrating imported devices into other kernel
   subsystems (human interface subsystem, audio subsystem, etc)
-\item Addition of ``NCD-admin'' utilities for managing and administering
+\item<8-> Addition of ``NCD-admin'' utilities for managing and administering
   NCD system
-\item Addition of graceful connect and disconnect handling in real-time
+\item<9-> Addition of graceful connect and disconnect handling in real-time
 \end{itemize}
@@
 \begin{itemize}
-\item Divorces peripherals from the computers to which they are
+\item<1-> Divorces peripherals from the computers to which they are
   attached
-\item Allows simple creation of a ``Device Server''
-\item Extends the ``Everything Is A File'' -nix Philosophy
-\item Allows more efficient use of devices
-\item Allows devices to be networked in a generalized, uniform manner
+\item<2-> Allows simple creation of a ``Device Server''
+\item<3-> Extends the ``Everything Is A File'' -nix Philosophy
+\item<4-> Allows more efficient use of devices
+\item<5-> Allows devices to be networked in a generalized, uniform manner
 \end{itemize}
@@
 \begin{itemize}
-\item Does not work well when Character Device is not the exclusive
+\item<1-> Does not work well when Character Device is not the exclusive
   device interface
@@
   \end{itemize}
-\item Requires exported devices to be ``de-integrated'' from host system
-\item Requires QOS guarantees and high guilty network connection
+\item<2-> Requires exported devices to be ``de-integrated'' from host system
+\item<3-> Requires QOS guarantees and high guilty network connection
 \end{itemize}
@@
 \begin{itemize}
-\item Quick, cheap, and simple ad-hoc network KVM systems
-\item Remote, networked web-cams, scanners, printers, etc
-\item Basis of ``protocol''-over-Ethernet systems (USB-over-Ethernet,
+\item<1-> Quick, cheap, and simple ad-hoc network KVM systems
+\item<2-> Remote, networked web-cams, scanners, printers, etc
+\item<3-> Basis of ``protocol''-over-Ethernet systems (USB-over-Ethernet,
   SATA-over-Ethernet, Etc)
-\item Basis of redundant, ``cloud-like'' device pools
-\item Basis of peripheral cluster systems (GPU Clusters, RNG Clusters,
+\item<4-> Basis of redundant, ``cloud-like'' device pools
+\item<5-> Basis of peripheral cluster systems (GPU Clusters, RNG Clusters,
   Etc)
 \end{itemize}
+
+\end{frame}
+
+%---KVM Slide---%
+\begin{frame}{\bf KVM}
+
+Building a networked KVM system with NCDs:
+
+  \begin{center}
+    \includegraphics[height=0.6\paperheight,keepaspectratio]{kvm.png}
+  \end{center}
+
+\end{frame}
+
+%---Device Pool Slide---%
+\begin{frame}{\bf Random Number Generator Pool}
+
+Building a distributed RNG system with NCDs:
+
+  \begin{center}
+    \includegraphics[height=0.6\paperheight,keepaspectratio]{rng.png}
+  \end{center}
 
@@
 \begin{itemize}
-\item Add multi-device, multi-server support
-\item Add kernel sub-system handler layer (HID, etc)
-\item Add device protection and removal capabilities to server system
-\item Add advertisement protocol
-\item Consider removing networking code from client kernel module
+\item<1-> Add multi-device, multi-server support
+\item<2-> Add kernel sub-system handler layer (HID, etc)
+\item<3-> Add device protection and removal capabilities to server system
+\item<4-> Add advertisement protocol
+\item<5-> Consider removing networking code from client kernel module
 \end{itemize}
@@
 
+%---Conclusion Slide---%
+\begin{frame}{\bf Conclusion}
+
+Network Character Devices provide:
+
+\begin{itemize}
+\item<1-> A simple means to add network functionality to devices using
+  well understood file semantics and interfaces.
+\item<2-> The groundwork for cloud-like device pools and arrays
+\item<3-> The means to efficiently export devices from where they are not
+  need and import them where they can be used.
+\end{itemize}
+
+But to make them widely applicable:
+
+\begin{itemize}
+\item<4-> Device drivers must move back toward the sole use of the
+  device file based APIs
+\item<5-> Advertisement and management tools must be added
+\item<6-> The OS must not assume that it has exclusive access to any
+  device connected to the physical hardware on which it is running.
+\end{itemize}
+
+\end{frame}
+
 \section{Bibliography}
+
 %---Bibliography---%

### 2011-12-09 report/presentation.tex
presentation: print all references, handout mode for no overlay
presentation: add udev and ncadmin frames

@@
 
-\documentclass[xcolor=dvipsnames]{beamer}
+\documentclass[handout]{beamer}
 \usetheme{AnnArbor}
@@
 
-\begin{frame}{\texttt{udev}}
+\begin{frame}[t,fragile]{\texttt{udev}}
+  \begin{lstlisting}[gobble=4,title=no way to set permissions within kernel,language=]
+    ACTION=="add", SUBSYSTEM=="netchar", MODE="0666"
+  \end{lstlisting}
+  \begin{lstlisting}[gobble=4,title=cannot retroactively apply input attributes,language=]
+    $ udevadm info -qall -n /dev/input/event8
+    P: /devices/virtual/input/input8/event8
+    N: input/event8
+    E: DEVNAME=/dev/input/event8
+    E: DEVPATH=/devices/virtual/input/input8/event8
+    E: ID_INPUT=1
+    E: ID_INPUT_KEY=1
+    E: SUBSYSTEM=input
+  \end{lstlisting}
 \end{frame}
 
-\begin{frame}{\texttt{ncadmin}}
+\begin{frame}[fragile]{\texttt{ncadmin}}
+  \begin{lstlisting}[gobble=4,language=]
+    usage: ncadmin command [args]
+
+      The following commands are supported:
+
+      add server port [name]
+        server - an ip address
+        port   - a port
+        name   - optional name relative to /dev. if no name is
+                 specified /dev/netchar/import# will be used
+
+      rm name
+        name   - the device relative to /dev to remove
+  \end{lstlisting}
+  \begin{itemize}
+    \item<2-> does nothing!
+    \item<3-> could use discovery service
+    \item<4-> requires individual \texttt{cdev}s
+    \item<5-> name is optional
+  \end{itemize}
 \end{frame}
@@
 %---Bibliography---%
-\begin{frame}{\bf Bibliography}
-
+\begin{frame}[t,allowframebreaks]{\bf Bibliography}
+\nocite{*}
 \bibliography{refs}
-
 \end{frame}

### 2011-12-10 report/presentation.tex
presentation: kill trailing whitespace
presentation: introduction and related work
Added goals page

@@
   University of Colorado at Boulder \\
-  \texttt{\{first.last\}@colorado.edu}
+  \texttt{matthew.monaco@colorado.edu} \\
+  \texttt{andrew.sayler@colorado.edu} \\
+  \texttt{landon.spear@colorado.edu}
 }
@@
 %---Intro Slide---%
-\begin{frame}{\bf NCD Overview}
-  We did some stuff, we used some references \cite{ldd3}.
-\end{frame}
 
@@
 
-\section{Related Work}
+\begin{frame}{\bf NCD Goals}
+  \begin{itemize}
+    \item<1-> Export any character device (or file) from a server over IP
+    \begin{itemize}
+      \item<2-> Client has exclusive access
+      \item<3-> User-level advertisment, kernel-level functionality
+    \end{itemize}
+    \item<4-> Non-locality should be transparent to client's user level operations    
+    \begin{itemize}
+      \item<5-> Imported devices are accessed as local devices
+      \item<6-> All client device accesses are intercepted by kernel module and sent over network
+      \item<7-> Server has no access so write conflicts are not a problem
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{\bf NCD Overview}
+  \begin{itemize}
+    \item<1-> Stuff
+    \item<2-> Motivation: transparent remote device usage
+    \item<3-> Client: kernel module (Linux 2.6.x+)
+    \item<4-> Server: userspace daemon
+    \item<5-> Current Project Status:
+    \begin{itemize}
+      \item<6-> successful implementation
+      \item<6-> rough around the edges
+      \item<6-> very clear focus for future work
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Related Work}
+  \begin{itemize}
+    \item nfs
+    \item nbd
+    \item CUSE
+    \item USB/IP Project
+  \end{itemize}
+\end{frame}
 

### 2011-12-10 report/presentation.tex
Goal page & Overview page

@@
   \begin{itemize}
-    \item<1-> Stuff
-    \item<2-> Motivation: transparent remote device usage
-    \item<3-> Client: kernel module (Linux 2.6.x+)
-    \item<4-> Server: userspace daemon
-    \item<5-> Current Project Status:
+    \item<1-> Motivation: providing transparent access to remote devices opens up a new paradigm of storage and device usage
     \begin{itemize}
-      \item<6-> successful implementation
-      \item<6-> rough around the edges
-      \item<6-> very clear focus for future work
+      \item<2-> Physical locality only needed for lower latency
+    \end{itemize}
+    \item<3-> Project attributes:
+    \begin{itemize}
+      \item<4-> Client: kernel module (Linux 2.6.x+)
+      \item<5-> Server: userspace daemon
+      \item<6-> Current Project Status:
+        \begin{itemize}
+          \item<7-> Successful implementation
+          \item<8-> Rough around the edges
+          \item<9-> More development required for certain devices
+          \item<10-> Very clear focus for future work
+        \end{itemize}
       \end{itemize}

### 2011-12-10 report/paper.tex
paper: abstract

@@
 \begin{abstract}
-Lorem ipsum
+While there are various projects which provide a network abstraction layer for
+specific high-level subsystems in the Linux kernel, there aren't any low-level
+general purpose solutions. Network character devices (NCD) is a system for
+exporting character device drivers of a fast, local network. However NCD works
+to some extent with regular files, block devices, sockets, and pipes; it is also
+suitably responsive over Internet connections for many applications.
 \end{abstract}
@@
 \end{document}
+
+% vim: set ts=2 sts=2 sw=2 et spell tw=72 : %

### 2011-12-10 report/paper.tex
paper: architecture section

@@
 
-\lstinputlisting[float=*,
-  caption={Listing Code Sample},
-  label=lst:codeSample.c]{codesample.c}
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=0.5\textwidth]{arch-03.png}
+  \end{center}
+  \caption{The basic client-server model of NCD.}
+  \label{fig:architecture}
+\end{figure}
 
-\section{System}
-\label{sec:system}
+On a very high level, NCD uses a basic client server model. The client
+issues a request to the server, which in turn issues a reply. The
+server never initiates any communication on its own.
+
+\subsection{Client}
+
+The client is currently implemented entirely within an out-of-tree kernel
+module. This is arguably not the most robust, nor secure, solution. See
+section \ref{sec:futurework} for details on splitting out some functionality
+into a user-space helper.
+
+Devices which have been imported on the client reside within the
+devtmpfs filesystem. This is almost always mounted on \texttt{/dev}. By
+default an imported device will take on the name
+\texttt{/dev/netchar/importN}, where \texttt{N} is the next available
+number. In the future (and without much effort), imported devices will
+be able to take on their original name from the server. For example, an
+input device can be located at \texttt{/dev/input/eventN}. Furthermore,
+the user or administrator may choose any mapping (s)he see fits.
+
+\subsection{Server}
+
+The server is, as mentioned, a typical user-space daemon. It
+currently only services a single file. Whether or not the server should
+handle multiple exports is still up for evaluation.
+
+The server listens on a port specified on the command line. There is no
+default. When a request is received from the client, it will correspond
+to one of the standard VFS file operations. Currently, \texttt{open()},
+\texttt{close()}
+\footnote{The counterpart for \texttt{close()} is actually referred to
+as \texttt{release()} within the kernel.}
+, \texttt{read()}, and \texttt{write()} are supported.
+
+After receiving a request, the server then performs the requested
+operation locally and returns the result to the client in a reply
+message. For example, if the client sends an \texttt{open()} request,
+the server will open and hold a file descriptor on the client's behalf.
+
+This straightforward architecture will actually support any file type
+(with certain restrictions for each type). Character devices, regular
+files, sockets, and pipes should work fully (when all file operations
+are implemented). Block devices and even directories will work to some
+extent.
+
+\subsection{Protocol}
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=0.5\textwidth]{proto.png}
+  \end{center}
+  \caption{NCD operates over a simple request reply protocol. Bulk data
+  immediately follows formal messages, as required.}
+  \label{fig:protocol}
+\end{figure}
+
+The protocol is implemented as a simple request-reply system over
+\texttt{INET} sockets. All request types are encapsulated within a
+single request structure (see section \ref{sec:implementation}).
+Similarly, there is only a single reply structure to account for all
+reply types.
+
+Bulk data is sent in the raw immediately following the corresponding
+request/reply. In the case of a \texttt{read()}, the targeted data
+immediately follows the reply. In the case of a \texttt{write()}, the
+targeted data immediately follows the initial request.
 

### 2011-12-10 report/presentation.tex
presentation: client/server swap (typo)
Related Work slide

@@
   \begin{itemize}
-    \item nfs
-    \item nbd
+    \item NFS
+    \begin{itemize}
+      \item Access to remote file system, but no devices
+      \item Safe access
+      \item http://nfs.sourceforge.net/
+    \end{itemize}
+    \item NBD
+    \begin{itemize}
+      \item Network Block Device- Introduced in Linux kernel 2.1.101
+      \item Works with any file system but no exclusive access 
+      \item http://nbd.sourceforge.net/
+    \end{itemize}
     \item CUSE
+    \begin{itemize}
+      \item Character Device in User Space (extension of FUSE)- Linux kernel 2.6.31
+      \item http://lwn.net/Articles/308445/
+    \end{itemize}
     \item USB/IP Project
+    \begin{itemize}
+      \item USB device sharing over IP
+      \item http://usbip.sourceforge.net/
+    \end{itemize}
+
   \end{itemize}
@@
   \begin{itemize}
-    \item<1-> The server side resides entirely in a single kernel module
-    \item<2-> The client side is a standard userspace server
+    \item<1-> The client side resides entirely in a single kernel module
+    \item<2-> The server side is a standard userspace daemon
   \end{itemize}

### 2011-12-10 report/presentation.tex
added more to intro slides

@@
     \begin{itemize}
-      \item<2-> Client has exclusive access
-      \item<3-> User-level advertisment, kernel-level functionality
+      \item<2-> Mouse, Webcam, keyboard, USB flash drive, Xbox Kinect, monitor, external HD, etc...
+      \item<3-> Client has exclusive access
+      \item<4-> User-level advertisment, kernel-level functionality
     \end{itemize}
-    \item<4-> Non-locality should be transparent to client's user level operations    
+    \item<5-> Non-locality should be transparent to client's user level operations    
     \begin{itemize}
-      \item<5-> Imported devices are accessed as local devices
-      \item<6-> All client device accesses are intercepted by kernel module and sent over network
-      \item<7-> Server has no access so write conflicts are not a problem
+      \item<6-> Imported devices are accessed as local devices
+      \item<7-> All client device accesses are intercepted by kernel module and sent over network
+      \item<8-> Server user-space daemon accesses device to emulate client
+      \item<9-> Server should have no other access so write conflicts are not a problem
     \end{itemize}
@@
           \item<9-> More development required for certain devices
-          \item<10-> Very clear focus for future work
+          \item<10-> Groundwork laid for exporting all devices
+          \item<11-> Very clear focus for future work
         \end{itemize}
@@
       \item Character Device in User Space (extension of FUSE)- Linux kernel 2.6.31
+      \item Used for OSS proxy for remote sound systems
       \item http://lwn.net/Articles/308445/

### 2011-12-10 report/presentation.tex
Added refrences for related work
Added refrences
Added copies of system overview diagrams and set toc depth to 1
Fixed minro types

@@
 \setbeamertemplate{navigation symbols}{}
+\setcounter{tocdepth}{1}
 
@@
 
+%---Full State Slide---%
+\begin{frame}[c]{Full System}
+  \begin{center}
+    \includegraphics[height=0.75\paperheight,keepaspectratio]{system-full.png}
+  \end{center}
+\end{frame}
+
+%---Current State Image Slide---%
+\begin{frame}[c]{Current State}
+  \begin{center}
+    \includegraphics[height=0.75\paperheight,keepaspectratio]{system-working.png}
+  \end{center}
+\end{frame}
+
 \begin{frame}{\bf NCD Overview}
   \begin{itemize}
-    \item<1-> Motivation: providing transparent access to remote devices opens up a new paradigm of storage and device usage
+    \item<1-> Motivation: providing transparent access to remote
+      devices opens up a new paradigm of storage and device usage
     \begin{itemize}
@@
   \begin{itemize}
-    \item NFS
+  \item NFS \cite{nfs-source}
     \begin{itemize}
@@
     \item Safe access
-      \item http://nfs.sourceforge.net/
     \end{itemize}
-    \item NBD
+  \item NBD \cite{nbd-source}
     \begin{itemize}
@@
     \item Works with any file system but no exclusive access 
-      \item http://nbd.sourceforge.net/
     \end{itemize}
-    \item CUSE
+  \item CUSE \cite{cuse-article}
     \begin{itemize}
-      \item Character Device in User Space (extension of FUSE)- Linux kernel 2.6.31
+    \item Character Device in User Space (extension of FUSE)
     \item Used for OSS proxy for remote sound systems
-      \item http://lwn.net/Articles/308445/
     \end{itemize}
-    \item USB/IP Project
+  \item USB/IP Project \cite{usbip-source}
     \begin{itemize}
     \item USB device sharing over IP
-      \item http://usbip.sourceforge.net/
+    \end{itemize}
+  \item Plan 9 \cite{plan9-pike}
+    \begin{itemize}
+    \item Full networked file based OS
+    \item Allows sharing devices over a network via file interfaces
+    \item Not widely deployed
+    \end{itemize}
   \end{itemize}
 
-  \end{itemize}
 \end{frame}
@@
   \end{lstlisting}
+%$
 \end{frame}
@@
 %---Full State Slide---%
-
 \begin{frame}[c]{Full System}
@@
     creation on client
-  \item<5-> Tested with various ``real'' and device files
+  \item<5-> Tested with various standard files and device files
   \end{itemize}
@@
 \item<2-> Requires exported devices to be ``de-integrated'' from host system
-\item<3-> Requires QOS guarantees and high guilty network connection
+\item<3-> Requires QOS guarantees and high quality network connection
 \end{itemize}
@@
 \item<3-> The means to efficiently export devices from where they are not
-  need and import them where they can be used.
+  needed and import them where they can be used.
 \end{itemize}

### 2011-12-11 report/presentation.tex
report: put listings in their own files
I should have done this from the start but I was lazy. Now I want to
reuse them in the paper. This is another form of laziness =)

@@
   breakatwhitespace=false,
-  title=\lstname,
   frame=single,
@@
 
-\begin{frame}[fragile]{Protocol}
+\begin{frame}{Protocol}
   \begin{columns}
     \begin{column}{0.5\textwidth}
-      \begin{lstlisting}[gobble=8]
-        enum FOPS {
-          FOP_OPEN, FOP_RELEASE,
-          FOP_READ, FOP_WRITE,
-          _FOP_ERROR
-        };
-
-        struct fop_request {
-
-          enum FOPS call;
-
-          union {
-            struct {
-              int     flags;
-              mode_t  mode;
-            };
-
-            size_t  count;
-          };
-        };
-      \end{lstlisting}
+      \lstinputlisting[lastline=19]{protocol.c}
     \end{column}
     \begin{column}{0.5\textwidth}
-      \begin{lstlisting}[gobble=8]
-        struct fop_reply {
-
-          enum FOPS call;
-
-          union {
-            int      open;
-            int      close;
-            ssize_t  read;
-            ssize_t  write;
-          };
-        };
-      \end{lstlisting}
+      \lstinputlisting[firstline=21]{protocol.c}
     \end{column}
@@
 
-\begin{frame}[fragile]{Hello, World!}
+\begin{frame}{Hello, World!}
   \begin{columns}
     \begin{column}{0.6\textwidth}
-      \begin{lstlisting}[gobble=8]
-        #include <linux/kernel.h>
-        #include <linux/module.h>
-
-        static int __init nc_init(void)
-        {
-          /* do stuff */
-          return 0;
-        }
-
-        static int __exit nc_exit(void)
-        {
-          /* undo stuff */
-        }
-
-        MODULE_LICENSE("GPL");
-
-        module_init(nc_init);
-        module_exit(nc_exit);
-
-      \end{lstlisting}
+      \lstinputlisting[]{kmod-hello.c}
     \end{column}
@@
 
-\begin{frame}[fragile]{Character Device Driver}
-  \begin{lstlisting}[gobble=4]
-    #include <linux/cdev.h>
-    #include <linux/device.h>
-    #include <linux/fs.h>
-
-    static dev_t          nc_dev_t;
-    static struct class*  nc_class;
-    static struct cdev*   nc_cdev;
-    static struct device* nc_device;
-
-  \end{lstlisting}
+\begin{frame}{Character Device Driver}
+  \lstinputlisting[]{cdev-driver.c}
 
@@
 
-\begin{frame}[fragile]{Filesystem Handler}
-  \begin{lstlisting}[gobble=4]
-    #include <linux/fs.h>
-
-    static int netchar_open(struct inode* inp, struct file* fp);
-    static ssize_t netchar_read(struct file* fp, char* buffer,
-                                size_t length, loff_t* offset);
-
-    static struct file_operations nc_fops = {
-      /* ... */
-      .open  = nc_open,
-      .read  = nc_read,
-    }
-  \end{lstlisting}
+\begin{frame}{Filesystem Handler}
+  \lstinputlisting[]{fop-driver.c}
 
@@
 
-\begin{frame}[fragile]{Server (\texttt{read()})}
-  \begin{lstlisting}[gobble=4]
-    struct fop_request req;
-    struct fop_reply   rep;
-    while (1) {
-      err = read(sd, &req, sizeof(req));
-      switch(req.call) {
-      case FOP_READ:
-        payload = realloc(payload, req.count);
-        rep.read = read(fd, payload, req.count);
-        if (rep.read < 0)
-          rep.read = -errno;
-        break;
-      }
-      err = write(sd, &rep, sizeof(rep));
-      if (rep.call == FOP_READ && rep.read > 0)
-        err = write(sd, payload, rep.read)
-    }
-  \end{lstlisting}
+\begin{frame}{Server (\texttt{read()})}
+  \lstinputlisting[]{server.c}
 \end{frame}
@@
 
-\begin{frame}[t,fragile]{\texttt{udev}}
-  \begin{lstlisting}[gobble=4,title=no way to set permissions within kernel,language=]
-    ACTION=="add", SUBSYSTEM=="netchar", MODE="0666"
-  \end{lstlisting}
-  \begin{lstlisting}[gobble=4,title=cannot retroactively apply input attributes,language=]
-    $ udevadm info -qall -n /dev/input/event8
-    P: /devices/virtual/input/input8/event8
-    N: input/event8
-    E: DEVNAME=/dev/input/event8
-    E: DEVPATH=/devices/virtual/input/input8/event8
-    E: ID_INPUT=1
-    E: ID_INPUT_KEY=1
-    E: SUBSYSTEM=input
-  \end{lstlisting}
+\begin{frame}[t]{\texttt{udev}}
+  \lstinputlisting[
+    lastline=1,
+    language={},
+    title=no way to set permissions within kernel,
+  ]{udev.txt}
+  \lstinputlisting[
+    firstline=2,
+    language={},
+    title=cannot retroactively apply input attributes
+  ]{udev.txt}
 %$
@@
 
-\begin{frame}[fragile]{\texttt{ncadmin}}
-  \begin{lstlisting}[gobble=4,language=]
-    usage: ncadmin command [args]
-
-      The following commands are supported:
-
-      add server port [name]
-        server - an ip address
-        port   - a port
-        name   - optional name relative to /dev. if no name is
-                 specified /dev/netchar/import# will be used
-
-      rm name
-        name   - the device relative to /dev to remove
-  \end{lstlisting}
+\begin{frame}{\texttt{ncadmin}}
+  \lstinputlisting[]{ncadmin.txt}
   \begin{itemize}

### 2011-12-11 report/inc/cdev-driver.c

@@ -1,0 +1,8 @@
+#include <linux/cdev.h>
+#include <linux/device.h>
+#include <linux/fs.h>
+
+static dev_t          nc_dev_t;
+static struct class*  nc_class;
+static struct cdev*   nc_cdev;
+static struct device* nc_device;

### 2011-12-11 report/inc/codesample.c

@@ -1,10 +1,0 @@
-/* This is a Listings test */
-int main(){
-
-	int i = 0;
-	for(i=0; i<10; i++){
-		printf("This is a LaTeX test");
-	}
-
-	return 0;
-}

### 2011-12-11 report/inc/fop-driver.c

@@ -1,0 +1,11 @@
+#include <linux/fs.h>
+
+static int netchar_open(struct inode* inp, struct file* fp);
+static ssize_t netchar_read(struct file* fp, char* buffer,
+size_t length, loff_t* offset);
+
+static struct file_operations nc_fops = {
+	/* ... */
+	.open  = nc_open,
+	.read  = nc_read,
+}

### 2011-12-11 report/inc/kmod-hello.c

@@ -1,0 +1,18 @@
+#include <linux/kernel.h>
+#include <linux/module.h>
+
+static int __init nc_init(void)
+{
+	/* do stuff */
+	return 0;
+}
+
+static int __exit nc_exit(void)
+{
+	/* undo stuff */
+}
+
+MODULE_LICENSE("GPL");
+
+module_init(nc_init);
+module_exit(nc_exit);

### 2011-12-11 report/inc/ncadmin.c

@@ -1,0 +1,10 @@
+ACTION=="add", SUBSYSTEM=="netchar", MODE="0666"
+
+$ udevadm info -qall -n /dev/input/event8
+P: /devices/virtual/input/input8/event8
+N: input/event8
+E: DEVNAME=/dev/input/event8
+E: DEVPATH=/devices/virtual/input/input8/event8
+E: ID_INPUT=1
+E: ID_INPUT_KEY=1
+E: SUBSYSTEM=input

### 2011-12-11 report/inc/protocol.c

@@ -1,0 +1,31 @@
+enum FOPS {
+	FOP_OPEN, FOP_RELEASE,
+	FOP_READ, FOP_WRITE,
+	_FOP_ERROR
+};
+
+struct fop_request {
+	
+	enum FOPS call;
+
+	union {
+		struct {
+			int flags;
+			mode_t mode;
+		};
+	
+		size_t count;
+	};
+};
+
+struct fop_reply {
+
+	enum FOPS call;
+
+	union {
+		int     open;
+		int     close;
+		ssize_t read;
+		ssize_t write;
+	};
+};

### 2011-12-11 report/inc/server.c

@@ -1,0 +1,16 @@
+struct fop_request req;
+struct fop_reply   rep;
+while (1) {
+	err = read(sd, &req, sizeof(req));
+	switch(req.call) {
+	case FOP_READ:
+		payload = realloc(payload, req.count);
+		rep.read = read(fd, payload, req.count);
+		if (rep.read < 0)
+			rep.read = -errno;
+		break;
+	}
+	err = write(sd, &rep, sizeof(rep));
+	if (rep.call == FOP_READ && rep.read > 0)
+		err = write(sd, payload, rep.read)
+}

### 2011-12-11 report/paper.tex
paper: convert existing figures to grayscale
paper: make listings more suitable this medium

@@
   language=C,
-  basicstyle=\footnotesize,
+  basicstyle=\tiny,
   numbers=none,
@@
   breakatwhitespace=false,
-  title=\lstname,
-  frame=single,
-  frameround=tttt
+  float
 }
@@
   \begin{center}
-    \includegraphics[width=0.5\textwidth]{arch-03.png}
+    \includegraphics[width=0.5\textwidth]{arch-03-g.png}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=0.5\textwidth]{proto.png}
+    \includegraphics[width=0.5\textwidth]{proto-g.png}
   \end{center}
@@
 \label{sec:implementation}
+
+\subsection{Protocol}
+
+\subsection{Client}
+
+\subsection{Misc}
+
+\subsubsection{udev}
+
+\subsubsection{ncadmin}
 

### 2011-12-11 report/paper.tex
paper: protocol implementation

@@
 
+\lstinputlisting[
+  caption={The structures which make up the NCD protocol.},
+  label=lst:protocol
+]{protocol.c}
+
+The protocol for NCD operates over TCP sockets. Despite the potential
+for large amounts of streaming data over the internet --- data which may
+be acceptable to drop --- some messages are unacceptable to lose. Any
+request or reply should be guaranteed. The data following a
+\texttt{read()} or \texttt{write()} (figure \ref{fig:protocol}), for
+example, may not be as important. Therefore, in the future it may be
+beneficial to implement the core protocol over TCP and bulk data over
+UDP (section \ref{sec:futurework}).
+
+Because there is only one type of packet for a request, and another for
+a reply, (listing \ref{lst:protocol}) there is a small amount of wasted
+space on some file operations. We have done our best to reduce this
+overhead as much as possible, making use of a cascade of nested unions
+and structs.
+
+The request struct is suitable to handle all of the function arguments
+for all of the operations we support. The reply struct is much simpler
+as functions only have a single return value. While some file operations
+have output parameters and some userspace system calls do as well, there
+aren't any cases where these need to be passed over the network.
+
 \subsection{Client}

### 2011-12-11 report/inc/cdev-driver.c
report: combine cdev-driver and fop-driver example

@@ -6,3 +6,13 @@ static dev_t          nc_dev_t;
 static struct class*  nc_class;
 static struct cdev*   nc_cdev;
 static struct device* nc_device;
+
+static int netchar_open(struct inode* inp, struct file* fp);
+static ssize_t netchar_read(struct file* fp, char* buffer,
+size_t length, loff_t* offset);
+
+static struct file_operations nc_fops = {
+	/* ... */
+	.open  = nc_open,
+	.read  = nc_read,
+}

### 2011-12-11 report/inc/fop-driver.c

@@ -1,11 +1,0 @@
-#include <linux/fs.h>
-
-static int netchar_open(struct inode* inp, struct file* fp);
-static ssize_t netchar_read(struct file* fp, char* buffer,
-size_t length, loff_t* offset);
-
-static struct file_operations nc_fops = {
-	/* ... */
-	.open  = nc_open,
-	.read  = nc_read,
-}

### 2011-12-11 report/paper.tex
paper: client implementation section

@@
 
+\lstinputlisting[
+  caption={The NCD kernel module has a number of roles.},
+  label=lst:client
+]{cdev-driver.c}
+
+The client resides entirely within a kernel module \footnote{The merits
+of including everything within a kernel module are discussed in section
+\ref{sec:currentstatus}.}. A typical userspace program can be sloppy
+with cleanup because all memory is freed on termination. In contrast, a
+kernel module must be very careful because its lifetime is shorter than
+the kernel itself. Therefore it is very import to correctly handle
+errors and cleanup all memory in the \texttt{module\_init} and
+\texttt{module\_exit} routines.
+
+Some of the basic features of the NCD kernel module are listed in
+listing \ref{lst:client}. Firstly, NCD dynamically requests a major and
+minor number from the running kernel \footnote{Statically assigned
+major/minor numbers are no longer required and are discouraged for new
+modules.}. These numbers are stored in a \texttt{struct dev\_t}. Once
+these numbers are assigned, the kernel will pass all file operations on
+character devices with the NCD major/minor number to the NCD module.
+
+The \texttt{struct class} is somewhat synonymous with a kernel
+subsystem. Because the major/minor number is dynamic, it is with this
+class that userspace is able to persistently refer to NCD character
+device nodes (see \ref{subsec:udev}).
+
+\texttt{Struct file\_operations} is used to define the handlers for all
+file operations which we support for our character devices \footnote{See
+\texttt{linux/fs.h} for a complete list of file operations.}.
+
+Most importantly, \texttt{struct cdev} ties together the major/minor
+number with the given set of file operations. With this it is possible
+for a kernel module to register for many major/minor numbers (typically
+a contiguous range), while having fine-grained control over which
+subsets are handled by which distinct groups of file operations. This
+will be important in the future as specific subsystems are handled
+differently by NCD to increase functionality.
+
+Finally, \texttt{struct device} is used to actually create a character
+device node under \texttt{/dev} in userspace. Unfortunately, not
+everything we might want to do to this device can be accomplished in the
+kernel (permissions for example).
+
 \subsection{Misc}
@@
 \subsubsection{udev}
+\label{subsec:udev}
 

### 2011-12-11 report/paper.tex
paper: udev subsection

@@
 
+\lstinputlisting[
+  lastline=1,
+  caption={A simple \texttt{udev} rule for managing NCD device
+  permissions.},
+  label=lst:udevrule
+]{udev.txt}
+
+A lot of magic is performed on character devices in userspace by
+\texttt{udev}. Currently we are using \texttt{udev} to manage the
+permissions of imported devices (listing \ref{lst:udevrule}).
+
+\lstinputlisting[
+  firstline=3,
+  caption={\texttt{udev} attributes required by userspace programs such
+  as the \texttt{Xserver} for knowing a device is a keyboard.},
+  label=lst:udevattr
+]{udev.txt}
+
+For more functionality, imported character devices must have certain
+attributes such as those for the keyboard in listing \ref{lst:udevattr}.
+Unfortunately setting them retroactively does not convince all programs
+of a particular devices type. Therefore interfacing with specific kernel
+subsystems such as \texttt{input} and \texttt{usb} will be required in
+the future for enhanced functionality.
+
 \subsubsection{ncadmin}

### 2011-12-11 report/paper.tex
paper: ncadmin subsection

@@
 
+\lstinputlisting[
+  caption={The potential management interface for the NCD system.},
+  label=lst:ncadmin
+]{ncadmin.txt}
+
+As NCD grows, it will be important for the client to be able to handle
+dynamically added and removed devices. Listing \ref{lst:ncadmin}
+provides an idea of what an interface would look like to provide such
+functionality. This would give NCD the ability to create its character
+devices anywhere under \texttt{/dev}; they could be placed in a similar
+location to the corresponding device on the server.
+
 % Andy Sections

### 2011-12-11 report/paper.tex
paper: server section

@@
 
+\subsection{Server}
+
+\lstinputlisting[
+  caption={A code snippet from the server illustrating a \texttt{read()}
+  file operation},
+  label=lst:server
+]{server.c}
+
+The server will spend most of its time sitting on a blocked read of the
+socket file descriptor. Once the client sends a request, it will be
+available to the server in the appropriate structure.
+
+In the snippet given in listing \ref{lst:server}, a \texttt{read()} file
+operation is handled. A buffer must be allocated with a size
+indicated by the original request. Next, the server can perform the
+corresponding operation and send a reply message to the client. Finally,
+the bulk data is sent to the client. All buffering is performed by the
+underlying \texttt{SOCK\_STREAM} and TCP protocols.
+
+The server itself is, and will always be, simple compared to the
+in-kernel module. As NCD grows, the server's functionality should remain
+fairly static. Section \ref{sec:futurework} discusses some enhancements
+which will make the server more stable, however it doesn't require many
+new features.
+
 \subsection{Misc}

### 2011-12-11 report/inc/ncadmin.c
report: inc/ncadmin.c was erroneous, remove it

@@ -1,10 +1,0 @@
-ACTION=="add", SUBSYSTEM=="netchar", MODE="0666"
-
-$ udevadm info -qall -n /dev/input/event8
-P: /devices/virtual/input/input8/event8
-N: input/event8
-E: DEVNAME=/dev/input/event8
-E: DEVPATH=/devices/virtual/input/input8/event8
-E: ID_INPUT=1
-E: ID_INPUT_KEY=1
-E: SUBSYSTEM=input

### 2011-12-11 report/paper.tex
paper: remove language=C from the ncadmin listing
paper: nocite{*}
Modified abstract

@@
   breakatwhitespace=false,
+  frame=none,
+  frameround=tttt,
   float
@@
 \title{Networked Character Devices}
-\author{Monaco, Matthew \and Sayler, Andrew \and Spear, Landon
-  \\ \and University of Colorado}
+\author{Monaco, Matthew \and Sayler, Andrew \and Spear, Landon \\
+  \and University of Colorado \\
+  \texttt{first.last@colorado.edu}
+}
 \date{\today}
@@
 general purpose solutions. Network character devices (NCD) is a system for
-exporting character device drivers of a fast, local network. However NCD works
-to some extent with regular files, block devices, sockets, and pipes; it is also
-suitably responsive over Internet connections for many applications.
+exporting character device interfaces over a network. This capability
+divorces the traditional requirement that a device can only be used by
+the host to which it is physically connected in favor of a model that
+allows a device to be connected to one host but used by another.
+In addition to networking the character device interface for which it
+is optimized, NCD also works with regular files, block devices,
+sockets, and pipes. We demonstrate how our NCD system works, as well
+as a few examples of the unique systems that can be built atop the NCD
+interface.
 \end{abstract}
@@
 \lstinputlisting[
+  language={},
   caption={The potential management interface for the NCD system.},
@@
 
+\nocite{*}
 \bibliography{refs}

### 2011-12-11 report/paper.tex
Minor formatting echnages, minor edits, and a few figures added

@@
 
-\usepackage[text={6in, 9in}, centering]{geometry}
+\usepackage[text={6.5in, 9in}, centering]{geometry}
 \usepackage{graphicx}
@@
   frame=none,
-  frameround=tttt,
-  float
+  frameround=tttt
 }
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.5\textwidth]{system-full-g.png}
+  \caption{The NCD system. See Figure \ref{fig:system-working}
+    for currently implemented system state.}
+  \label{fig:system-full}
+\end{figure}
+
 \section{Related Work}
@@
 \begin{figure}[h]
-  \begin{center}
+  \centering
   \includegraphics[width=0.5\textwidth]{arch-03-g.png}
-  \end{center}
   \caption{The basic client-server model of NCD.}
@@
 be able to take on their original name from the server. For example, an
-input device can be located at \texttt{/dev/input/eventN}. Furthermore,
+input device should be located at \texttt{/dev/input/eventN}. Furthermore,
 the user or administrator may choose any mapping (s)he see fits.
@@
 
-The server is, as mentioned, a typical user-space daemon. It
+The server is a typical user-space daemon. It
 currently only services a single file. Whether or not the server should
@@
 \begin{figure}[h]
-  \begin{center}
+  \centering
   \includegraphics[width=0.5\textwidth]{proto-g.png}
-  \end{center}
   \caption{NCD operates over a simple request reply protocol. Bulk data
@@
 Some of the basic features of the NCD kernel module are listed in
-listing \ref{lst:client}. Firstly, NCD dynamically requests a major and
+listing \ref{lst:client}. NCD dynamically requests a major and
 minor number from the running kernel \footnote{Statically assigned
@@
 number with the given set of file operations. With this it is possible
-for a kernel module to register for many major/minor numbers (typically
+for a kernel module to register many major/minor numbers (typically
 a contiguous range), while having fine-grained control over which
@@
 A lot of magic is performed on character devices in userspace by
-\texttt{udev}. Currently we are using \texttt{udev} to manage the
+the Linux \texttt{udev} system.
+Currently we are using \texttt{udev} to manage the
 permissions of imported devices (listing \ref{lst:udevrule}).
@@
 
-The Network Character Device system is still a work in
-progress. We will need to Expand the system beyond it's current state
+The Network Character Devices system is still a work in
+progress. We will need to expand the system beyond it's current state
 before the NCD concept is ready for production-level use.
 
-Non the less, we have gained many insights into the advantages such a
+None the less, we have gained many insights into the advantages such a
 system might provide. We have also encountered the difficulties
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.5\textwidth]{system-working-g.png}
+  \caption{The NCD system with currently functioning portions
+    highlighted.}
+  \label{fig:system-working}
+\end{figure}
+
 As of this writing, we have completed a partial implementation of
-the Network Character Device system. Current functionality includes:
+the Network Character Devices system. Current functionality includes:
 
@@
 \item Support for importing a single character device on the client
-\item Client-side kernel-space Linux module implementation
-\item Server-side user-space background process implementation
-\item Support for open, close, read, and write calls
-\item Basic Linux udev support for automatic node creation on client
+\item Support for \texttt{open()}, \texttt{close()}, \texttt{read()},
+  and \texttt{write()} calls
+\item Basic Linux \texttt{udev} support for automatic node
+  provisioning on the client
 \end{packed_item}
@@
 This set of core functionality provides a working proof-of-concept
-Network Character Device system. This system fully supports the
+Network Character Devices system. This system fully supports the
 export of a single character device from server to client over the
@@
   subsystems (human interface subsystem, audio subsystem, etc)
-\item Addition of ``NCD-admin'' utilities for managing and administering
-  NCD system
+\item Addition of UDP and TCP support with user selection
+\item Addition of graceful connect and disconnect handling in real-time
 \end{packed_item}

### 2011-12-11 report/paper.tex
Minro biblio chnages
Added images and advantages section

@@
 Devices which have been imported on the client reside within the
-devtmpfs filesystem. This is almost always mounted on \texttt{/dev}. By
+devtmpfs file system. This is almost always mounted on \texttt{/dev}. By
 default an imported device will take on the name
@@
 The protocol for NCD operates over TCP sockets. Despite the potential
-for large amounts of streaming data over the internet --- data which may
+for large amounts of streaming data over the Internet --- data which may
 be acceptable to drop --- some messages are unacceptable to lose. Any
@@
 
+The primary advantages of using Networked Charterer Devices are:
+
+\begin{packed_item}
+\item Divorces peripherals from the computers to which they are
+  attached
+\item Allows simple creation of a ``Device Server''
+\item Extends the ``Everything Is A File'' -nix philosophy
+\item Allows more efficient use of devices
+\item Allows devices to be networked in a generalized, uniform manner
+\end{packed_item}
+
+By allowing the interface for any device to be exported over the
+network, the NCD system effectively removes the traditional restriction
+that devices are only available on the hosts to which they are
+physically connected. Removing this restriction has far reaching
+implications (see Section \ref{sec:potentialuses}). One of which is
+increased efficiency of device utilization. By allowing devices to
+be exported from machines where they are not needed and imported on
+machines where they are, we can insure that devices are efficiently
+utilized without having to worry about the underlying physical locations
+of the devices.
+
+One of the foremost tenants of the Unix philosophy is the ``Everything
+is a file'' concept. The NCD system extends this concept to
+``Everything is a networked file''. This allows us to create ``Device
+Servers'' as easily as we currently create file servers. The NCD
+system allows us to network any device via the general and
+well-understood file interface instead of relying on devices specific
+networking implementations.
+
 \subsection{Challenges}
@@
 \label{sec:futurework}
+
+\section{Potential Uses}
+\label{sec:potentialuses}
+
+\subsection{Networked KVM System}
+\label{sec:networkedkvm}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.5\textwidth]{kvm-g.png}
+  \caption{A networked KVM system built using NCD.}
+  \label{fig:kvm}
+\end{figure}
+
+\subsection{Distributed RNG System}
+\label{sec:distributedrng}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.5\textwidth]{rng-g.png}
+  \caption{A distributed Random Number Generator system built using NCD.}
+  \label{fig:rng}
+\end{figure}
 

### 2011-12-11 report/paper.tex
Added Challenges section

@@
 
+While Network Charter Devices provide a number of advantages, we have
+also stumbled upon a number of challenges when it comes to properly
+building and utilizing an NCD system. In particular, the NCD system:
+
+\begin{packed_item}
+\item Does not work well when Character Device file is not the exclusive
+  device interface
+\item Requires exported devices to be ``de-integrated'' from host system
+\item Requires QOS guarantees and high quality network connection
+\end{packed_item}
+
+The original intention of the Unix and Linux \texttt{dev} file
+system and the corresponding Charter Device file interface was to
+provide the sole interface to device drivers. Today, while Character
+Device file interfaces are still is use, they are no longer the sole
+devise driver interface. Auxiliary systems like procfs
+\cite{interfaces}, sysfs \cite{sysfs-mochel}, and netlink sockets
+\cite{netlink-ayuso} provide supplementary and/or replacement interfaces
+for the tradition Character Device file interface. Furthermore, some
+dives interact directly with in-kernel subsystems (HID subsystem,
+etc). Since our system
+focuses on arbitrating the Charter Device file interface over the
+network, use of these additional or alternate interfaces make devices
+incompatible with the NCD system. This incompatibility can be mitigated
+by extending the NCD system to include these alternate interfaces, but
+breakdown of the original Unix ideal of having a dingle device driver
+interface makes building system like NCD much more difficult and less
+generalized.
+
+Most modern Operating Systems, including Linux, assume that any device
+physically connected to the system is available for the system to
+use. As mentioned previously, Network Character Devices break down this
+assumption. This one of the challenges involved in implementing an NCD
+system is to remove the OS's assumption that any connected devices in
+available for local use. Modern-day plug-and-play auto-configuration
+system make this task even more difficult, and the means for breaking
+down the availability assumption tends to be unique to a given
+OS. Effectively removing this assumption is one of the larger,
+still-open, NCD challenges.
+
+Finally, since the NCD system makes a networked device appear as
+though it is connect locally, the NCD system does not gracefully
+handle situations that would never occur locally like network faults
+and errors. Thus, the NCD system performs best when running on a
+network with QOS guarantees and the necessary throughput to support all
+NCD traffic.
+
 \section{Future Work}

### 2011-12-11 report/paper.tex
Fixed some typoes

@@
 \cite{netlink-ayuso} provide supplementary and/or replacement interfaces
-for the tradition Character Device file interface. Furthermore, some
-dives interact directly with in-kernel subsystems (HID subsystem,
-etc). Since our system
-focuses on arbitrating the Charter Device file interface over the
+for the traditional Character Device file interface. Furthermore, some
+device drivers interact directly with in-kernel subsystems (HID subsystem,
+etc), providing no direct userspace interface at all. Since our system
+focuses on abstracting the Charter Device file interface over the
 network, use of these additional or alternate interfaces make devices
 incompatible with the NCD system. This incompatibility can be mitigated
-by extending the NCD system to include these alternate interfaces, but
-breakdown of the original Unix ideal of having a dingle device driver
-interface makes building system like NCD much more difficult and less
+by extending the NCD system to include these alternate interfaces, but the
+breakdown of the original Unix ideal of having a single device driver
+interface makes building systems like NCD much more difficult and less
 generalized.
@@
 use. As mentioned previously, Network Character Devices break down this
-assumption. This one of the challenges involved in implementing an NCD
+assumption. Thus, one of the challenges involved in implementing an NCD
 system is to remove the OS's assumption that any connected devices in
 available for local use. Modern-day plug-and-play auto-configuration
-system make this task even more difficult, and the means for breaking
-down the availability assumption tends to be unique to a given
+systems make this task even more difficult, and the means for breaking
+down the availability assumption tend to be unique to a given
 OS. Effectively removing this assumption is one of the larger,
@@
 Finally, since the NCD system makes a networked device appear as
-though it is connect locally, the NCD system does not gracefully
+though it is connected locally, the NCD system does not gracefully
 handle situations that would never occur locally like network faults

### 2011-12-11 report/paper.tex
paper: I meant 'for'

@@
 number with the given set of file operations. With this it is possible
-for a kernel module to register many major/minor numbers (typically
+for a kernel module to register for many major/minor numbers (typically
 a contiguous range), while having fine-grained control over which

### 2011-12-11 report/paper.tex
Added future work section

--- 2011-12-11 report/paper.tex
@@
 As of this writing, we have completed a partial implementation of
-the Network Character Devices system. Current functionality includes:
+the Network Character Devices system (Figure
+\ref{fig:system-working}). Current functionality includes:
 
@@
 
+As Figure \ref{fig:system-working} shows, there are still several
+portions of the NCD system that need implemented before the system is
+ready for production-level use. In particular, the top priorities
+currently are:
+
+\begin{packed_item}
+\item Add multi-device, multi-server support
+\item Add kernel sub-system handler layer (HID, etc)
+\item Add device protection and removal capabilities to server system
+\item Add advertisement protocol
+\item Separate networking and memory-intensive code from client kernel module
+\end{packed_item}
+
+The current NCD system only support a single device per server and per
+client. To be truly useful and to enable the construction of the
+kinds of system we envision NCD being used for (Section
+\ref{potentialuses}), we must at least
+expand the client-side
+system to support importing an arbitrary number of devices from one
+or more servers. Ideally, servers should also be capable of serving
+multiple devices, but this is of less importance than multi-device
+client handling.
+
+As we mentioned in Section \ref{sec:challenges}, some modern device
+drivers bypass the traditional Character Device file interface. Thus, we
+must expand the NCD system to accommodate the additional interfaces and
+subsystems that are commonly in use. This includes special import
+handling for in-kernel subsystems like the X Server and HID
+subsystems, as well as generalized support for sysfs and procfs import
+and export.
+
+We also plan to add a means for device protection and exclusion to
+the NCD system. This will include both the need to remove exported
+devices from use by local server-side operating systems as well as controlling
+the number of clients that can connect to a single exported
+device. Controlling the latter point will require a per-device
+definition, probably provided as part of the export configuration, of
+how many clients may simultaneously utilize each device. Some devices will
+require exclusive access while other may be capable of handling read
+and/or write request from multiple clients simultaneously.
+
+Since large scale NCD deployments may include thousands of exported
+devices, the NCD system requires the addition of an advertisement and
+discovery protocol so that client may ascertain what resources are
+currently available for import. Such a system will need to provide
+information about what devices are available, as well as metadata
+describing the basic characteristics of a given device.
+
+Finally, the current client-side NCD implementation performs both direct
+network access and memory-intensive processing directly in the kernel
+code. For security reasons, it is desirable to separate the direct
+network communication from the kernel code to user-space code where
+the repercussions of a security breach as less extreme.
+For performance reasons, it is desirable to move memory
+intensive operations from the kernel code (where they are limited to
+the available kernel memory space) to user-space code where the memory
+availability is far greater and the use restrictions more relaxed. To
+solve these issues, we
+foresee building a client-daemon that will handle the bulk to the
+client-side operations and that will, in turn, communicate directly with
+a more minimal client kernel module only when necessary.
+
 \section{Potential Uses}

### 2011-12-11 report/paper.tex
Added Potnetial Uses and Conclusion, Fixed rng.svg server<->client mistake

@@
 kinds of system we envision NCD being used for (Section
-\ref{potentialuses}), we must at least
+\ref{sec:potentialuses}), we must at least
 expand the client-side
@@
 
+The shift from the traditional concept of locally available devices to a
+broader concept of generalized network accessible devices has the
+potential to allow the creation of a variety of novel
+applications. Here we explore a few of the possible applications of an
+NCD system.
+
 \subsection{Networked KVM System}
 \label{sec:networkedkvm}
+
+KVM (Keyboard-Video-Mouse) systems are commonly used in data centers
+and other areas where a user wishes to use a single keyboard, video,
+and mouse, to control multiple machines. KVM systems range from simple KVM
+switches that allow the user to select which local machine is
+physically connected to his or her i/o devices to
+networked KVM systems that allow the user to use a
+local KVM setup to interface with a computer located elsewhere on the
+network. The latter class of networked KVM devices traditionally require
+expensive propriety hardware.
 
@@
 
+As Figure \ref{fig:kvm} shows, we can use Networked Character Devices
+to quickly and easily build a networked KVM system in software,
+forgoing the need for expensive specialized hardware. The user simply
+connects a spare mouse and keyboard to an NCD server, tells the NCD
+system to export these devices, and then imports them on the client
+machine (s)he wishes to control. To the client, the devices appear as
+a locally connected keyboard and mouse, and the user may issue input to
+the client machine via the keyboard and mouse connected to the
+server.
+
+As we mentioned previously, an NCD based KVM system would require the
+appropriate import interface to the Linux HID subsystem, but once this is
+in place, building a networked KVM atop the NCD system would be a
+rather trivial matter.
+
 \subsection{Distributed RNG System}
 \label{sec:distributedrng}
+
+Another example of the novel systems that can be built use Network
+Charter Devices is a distributed random number generator. Linux has a
+fairly robust random number generator built into it and accessible as
+a Character Device at \texttt{/dev/random}. This RNG relies on the
+local entropy pool (generally filled by human interaction, interrupt
+timings, and other high entropy events) to generate random
+numbers. This, the speed at which random numbers can be generated is
+limited on any single system to the locally available
+entropy. This can severely hinder the
+performance of systems that require random numbers\footnote{Linux also
+provides the \texttt{/dev/urandom} RNG which is a pseudorandom number
+generator seeded off teh true
+\texttt{/dev/random}. \texttt{/dev/urandom} trades ``randomness'' in
+favor of speed. It is thus handy when one just needs to quickly
+generate pseudorandom numbers, but is not appropriate for cryptography or
+other secure random system.}.
 
@@
 
+We can imagine a system that uses Network Character Devices to overcome
+the normal limitations related to relying on locally available entropy. This
+system would be capable of generating cryptographically secure
+randomness while also delivering high performance.
+
+Figure \ref{fig:rng} shows how such a system might work. In this
+distributed RNG system, multiple hosts act as NCD servers by exporting
+their local \texttt{/dev/random} character devices. These devices are
+all imported by a single client running some from on RNG aggregator
+that combines the relatively slow output form the local RNGs into a
+single high speed RNG output. We can effectively pool the entropy
+available on numerous hosts to drive a single high performance RNG
+system.
+
+As in the KVM system, using Networked Character Devices makes the
+implementation of such a system relatively straight forward.
+
 \section{Conclusion}
 \label{sec:conclusion}
+
+Network Character Devices provide an extension of the standard device
+driver file interface across the network. They allow for device
+capabilities to be exported from where they are not needed and imported
+where they are. By separating device usage from the device physical
+location and host, the NCD system provides a generalized, transparent,
+means for abstracting device functionality across a network.
+
+As the potential uses demonstrate, the power of
+the NCD concept is in the ability to achieve complex solutions without
+needing to resort to specialized device networking protocols and other
+complex ad-hoc systems. We hope to continue the development of the NCD
+concept and to utilize it to build powerful, but easily implementable,
+device-dependent systems.
 

### 2011-12-12 report/paper.tex
Added intro and related work

@@
   \setlength{\parsep}{0pt}
+\label{sec:introduction}
+
+We did some stuff, we used some references \cite{ldd3}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.5\textwidth]{system-full-g.png}
+  \caption{The NCD system. See Figure \ref{fig:system-working}
+    for currently implemented system state.}
+  \label{fig:system-full}
+\end{figure}
+
+\section{Related Work}
+\label{sec:relatedwork}
+
 }{\end{itemize}}
@@
 
-We did some stuff, we used some references \cite{ldd3}.
+We have designed and implemented Network Character Devices (NCD) to
+enable a new approach to using and sharing devices. Today, networking 
+technology is becoming faster and more reliable at incedible rates. 
+More functionality that has been traditionally thought to only be 
+feasabily locally accessable can now be pushed to remote
+machines. Using client-server operation, our project aims to allow 
+a character device (or devices) present on a server to be exported to
+a client over a network connection. While character devices come in 
+many forms, from Webcams, mice and monitors to USB flash drives and 
+external hard drives, NCD works below where these devices are 
+interpreted and exports their device file over IP. While designed for
+character devices, this generality will actually allow us to export 
+any file, which in a *nix system means almost anything on the
+system. Being able to dynamically export devices over network
+connections suggests a new cloud-like paradigm of device usage where 
+locality is required only for reduced latency, which may not be a
+concern for certain devices.
+
+When a device is being exported by NCD we would like it to give the
+appearence of having essentially ``moved'' the device to the client 
+by forbidding user level accesses on the server. This is not part 
+of our current iteration of NCD, but it will be included as higher
+level functionality is implemented. Aquiring a remote device will be
+done through an advertisment protocol that exists in user space so
+that details may be supplied about the server and how the connection
+should be made (e.g. UDP vs TCP).  After this NCD's functionality on
+the client exists in kernel space so as to maintain transparency to
+user level operations. Once imported, a device should be accessed just
+like a normal local device.  This is acheived by the NCD kernel module
+intercepting accesses and sending them out over the network to the
+server. On the server, a userspace daemon accesses the physical
+device to emulate the client's activity. The overall structure of the
+NCD system is present in figure 1.
 
@@
 
+Currently we have achieved a successful implementation of the low
+level functionality for NCD. The client uses a kernel module written
+for linux 2.6.x+ while the server uses a userspace daemon. NCD is
+still rough and in need of further development, as it tends to be
+crash prone if not treated with care. The groundwork has been laid for
+exporting any device, however many device subsystems access their
+devices by avoiding the character device interface and these must be
+dealt with on an individual basis. Overall, the path for future work
+is very clear.
+
 \section{Related Work}
@@
 
-We are unique.
+There are many other projects in existence that implement similar
+network sharing in one form or another that provided inspiration for
+NCD.
+
+NFS was the first implementation of truely general information sharing
+over remote network connections.  However, NFS uses a specific file
+system and cannot interface directly with devices.\cite{nfs-source}
+
+NBD (Network Block Device) was included into the linux kernel starting
+with version 2.1.101. It enables users to export block devices over the
+network, allowing use of any filesystem. However, NBD provides no
+exclusive access guarantees and users must acertain on their own
+whether the remote drive is being used.\cite{nbd-source}
+
+CUSE (Character Device in User Space) is an extension of FUSE that
+allows bringing a characer device into a FUSE filesystem.  This allows
+users to access the device directly and potentially export it using
+user-level networking.\cite{cuse-article}
+
+USB/IP is a project to share USB devices over IP. This project is
+very similar to NCD (as USB devices are character devices) but is
+tailored specifinterceptingically to allow sharing USB devices over a
+network connection by intercepting USB specific data and encapsulating
+it in IP.\cite{usbip-source}
+
+Plan 9 is a distributed operating system developed as a successor to
+Unix by Bell Labs.  It implicitly allows sharing devices over a
+network via file interfaces. However, plan 9 does not currenty see
+wide deployment.\cite{plan9-pike}
 
@@
 % vim: set ts=2 sts=2 sw=2 et spell tw=72 : %
+intercepting

### 2011-12-12 report/paper.tex
paper: fix so it builds
paper: get the obvious typos. still haven't read through

@@
 \label{sec:introduction}
-
-We did some stuff, we used some references \cite{ldd3}.
 
@@
 level functionality for NCD. The client uses a kernel module written
-for linux 2.6.x+ while the server uses a userspace daemon. NCD is
+for Linux 2.6.x+ while the server uses a userspace daemon. NCD is
 still rough and in need of further development, as it tends to be
@@
 
-NFS was the first implementation of truely general information sharing
+NFS was the first implementation of truly general information sharing
 over remote network connections.  However, NFS uses a specific file
@@
 
-NBD (Network Block Device) was included into the linux kernel starting
+NBD (Network Block Device) was included into the Linux kernel starting
 with version 2.1.101. It enables users to export block devices over the
 network, allowing use of any filesystem. However, NBD provides no
-exclusive access guarantees and users must acertain on their own
+exclusive access guarantees and users must ascertain on their own
 whether the remote drive is being used.\cite{nbd-source}
@@
 CUSE (Character Device in User Space) is an extension of FUSE that
-allows bringing a characer device into a FUSE filesystem.  This allows
+allows bringing a character device into a FUSE filesystem.  This allows
 users to access the device directly and potentially export it using
@@
 very similar to NCD (as USB devices are character devices) but is
-tailored specifinterceptingically to allow sharing USB devices over a
+tailored specifically to allow sharing USB devices over a
 network connection by intercepting USB specific data and encapsulating
@@
 Unix by Bell Labs.  It implicitly allows sharing devices over a
-network via file interfaces. However, plan 9 does not currenty see
+network via file interfaces. However, plan 9 does not currently see
 wide deployment.\cite{plan9-pike}
@@
 % vim: set ts=2 sts=2 sw=2 et spell tw=72 : %
-intercepting

### 2011-12-12 report/paper.tex
Fixed paper.tex errors

@@
   \setlength{\parsep}{0pt}
-\label{sec:introduction}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.5\textwidth]{system-full-g.png}
-  \caption{The NCD system. See Figure \ref{fig:system-working}
-    for currently implemented system state.}
-  \label{fig:system-full}
-\end{figure}
-
-\section{Related Work}
-\label{sec:relatedwork}
-
 }{\end{itemize}}
@@
 device to emulate the client's activity. The overall structure of the
-NCD system is present in figure 1.
+NCD system is present in Figure \ref{fig:system-full}.
 
@@
 over remote network connections.  However, NFS uses a specific file
-system and cannot interface directly with devices.\cite{nfs-source}
+system and cannot interface directly with devices \cite{nfs-source}.
 
@@
 exclusive access guarantees and users must ascertain on their own
-whether the remote drive is being used.\cite{nbd-source}
+whether the remote drive is being used \cite{nbd-source}.
 
@@
 users to access the device directly and potentially export it using
-user-level networking.\cite{cuse-article}
+user-level networking \cite{cuse-article}.
 
@@
 network connection by intercepting USB specific data and encapsulating
-it in IP.\cite{usbip-source}
+it in IP \cite{usbip-source}.
 
@@
 Unix by Bell Labs.  It implicitly allows sharing devices over a
-network via file interfaces. However, plan 9 does not currently see
-wide deployment.\cite{plan9-pike}
+network via file interfaces. However, Plan 9 does not currently see
+wide deployment \cite{plan9-pike}.
 

### 2011-12-12 report/paper.tex
Fixed spelling errors

@@
 enable a new approach to using and sharing devices. Today, networking 
-technology is becoming faster and more reliable at incedible rates. 
+technology is becoming faster and more reliable at incredible rates. 
 More functionality that has been traditionally thought to only be 
-feasabily locally accessable can now be pushed to remote
+feasibly locally accessible can now be pushed to remote
 machines. Using client-server operation, our project aims to allow 
@@
 a client over a network connection. While character devices come in 
-many forms, from Webcams, mice and monitors to USB flash drives and 
+many forms, from Web cams, mice and monitors to USB flash drives and 
 external hard drives, NCD works below where these devices are 
@@
 When a device is being exported by NCD we would like it to give the
-appearence of having essentially ``moved'' the device to the client 
+appearance of having essentially ``moved'' the device to the client 
 by forbidding user level accesses on the server. This is not part 
 of our current iteration of NCD, but it will be included as higher
-level functionality is implemented. Aquiring a remote device will be
-done through an advertisment protocol that exists in user space so
+level functionality is implemented. Acquiring a remote device will be
+done through an advertisement protocol that exists in user space so
 that details may be supplied about the server and how the connection
@@
 user level operations. Once imported, a device should be accessed just
-like a normal local device.  This is acheived by the NCD kernel module
+like a normal local device.  This is achieved by the NCD kernel module
 intercepting accesses and sending them out over the network to the
@@
 provides the \texttt{/dev/urandom} RNG which is a pseudorandom number
-generator seeded off teh true
+generator seeded off the true
 \texttt{/dev/random}. \texttt{/dev/urandom} trades ``randomness'' in

### 2011-12-13 report/paper.tex
paper: make a pdf index, make references clickable
Added final changes and fixed hyperref version dependency

@@
 \usepackage{listings}
+\usepackage{hyperref}
 
@@
   frameround=tttt
+}
+
+\hypersetup{
+    colorlinks,
+    citecolor=black,
+    filecolor=black,
+    linkcolor=black,
+    urlcolor=black
 }
@@
 technology is becoming faster and more reliable at incredible rates. 
-More functionality that has been traditionally thought to only be 
-feasibly locally accessible can now be pushed to remote
-machines. Using client-server operation, our project aims to allow 
-a character device (or devices) present on a server to be exported to
-a client over a network connection. While character devices come in 
-many forms, from Web cams, mice and monitors to USB flash drives and 
-external hard drives, NCD works below where these devices are 
-interpreted and exports their device file over IP. While designed for
-character devices, this generality will actually allow us to export 
-any file, which in a *nix system means almost anything on the
-system. Being able to dynamically export devices over network
+Functionality that has been traditionally believed to only be 
+feasible at the local level can now be pushed to remote
+machines. Our project aims to allow 
+a character device present on a server to be exported to
+a client over a network connection. Many systems use character device
+interfaces, from Web cams, mice and monitors to USB flash drives and 
+external hard drives. NCD works by abstracting the character device interface 
+over IP. In Unix-like operating system, the character device interface
+is a file interface. While designed and optimized for
+character devices, this generality actually allows the NCD system to export 
+any file. Being able to dynamically export devices over network
 connections suggests a new cloud-like paradigm of device usage where 
@@
 
-When a device is being exported by NCD we would like it to give the
+When a device is exported via the NCD system we would like it to give the
 appearance of having essentially ``moved'' the device to the client 
-by forbidding user level accesses on the server. This is not part 
-of our current iteration of NCD, but it will be included as higher
-level functionality is implemented. Acquiring a remote device will be
+by forbidding user level accesses on the server.
+Acquiring a remote device will be
 done through an advertisement protocol that exists in user space so
 that details may be supplied about the server and how the connection
-should be made (e.g. UDP vs TCP).  After this NCD's functionality on
-the client exists in kernel space so as to maintain transparency to
+should be made (e.g. UDP vs TCP).  The client-side NCD's functionality
+exists in kernel space so as to maintain transparency to
 user level operations. Once imported, a device should be accessed just
 like a normal local device.  This is achieved by the NCD kernel module
-intercepting accesses and sending them out over the network to the
+intercepting i/o calls and sending them out over the network to the
 server. On the server, a userspace daemon accesses the physical
@@
 devices by avoiding the character device interface and these must be
-dealt with on an individual basis. Overall, the path for future work
-is very clear.
+dealt with on an individual basis. Section \ref{sec:futurework}
+details some of these challenges and our plans for their resolution.
 
@@
 
-NFS was the first implementation of truly general information sharing
-over remote network connections.  However, NFS uses a specific file
-system and cannot interface directly with devices \cite{nfs-source}.
+NFS was the first implementation of generalized file sharing
+over remote network connections.  However, NFS only supports normal
+files and cannot interface directly with devices \cite{nfs-source}.
 
-NBD (Network Block Device) was included into the Linux kernel starting
+NBD (Network Block Device) was rolled into the Linux kernel starting
 with version 2.1.101. It enables users to export block devices over the
-network, allowing use of any filesystem. However, NBD provides no
+network, allowing low level access to disks and other block devices.
+However, NBD provides no
 exclusive access guarantees and users must ascertain on their own
-whether the remote drive is being used \cite{nbd-source}.
+whether the remote drive is being used. It also mandates buffered
+operation and is optimized for block, not character, device operation
+\cite{nbd-source}.
 
@@
 users to access the device directly and potentially export it using
-user-level networking \cite{cuse-article}.
+user-level networking \cite{cuse-article}. We considered using CUSE in
+our NCD implementation, but opted for a direct in-kernel implementation
+instead since it is more standard.
 
-USB/IP is a project to share USB devices over IP. This project is
+USB/IP is a project to share USB devices over IP networks. This project is
 very similar to NCD (as USB devices are character devices) but is
@@
 network connection by intercepting USB specific data and encapsulating
-it in IP \cite{usbip-source}.
+it in IP \cite{usbip-source}. It does not support a more generalized
+notion of general character device networking.
 
@@
 network via file interfaces. However, Plan 9 does not currently see
-wide deployment \cite{plan9-pike}.
+wide deployment \cite{plan9-pike}. We hope to bring the NCD concept to
+the more widely deployed Linux environment.
 
@@
 module. This is arguably not the most robust, nor secure, solution. See
-section \ref{sec:futurework} for details on splitting out some functionality
+Section \ref{sec:futurework} for details on splitting out some functionality
 into a user-space helper.
@@
 The server is a typical user-space daemon. It
-currently only services a single file. Whether or not the server should
+currently only serves a single file. Whether or not the server should
 handle multiple exports is still up for evaluation.
@@
   \includegraphics[width=0.5\textwidth]{proto-g.png}
-  \caption{NCD operates over a simple request reply protocol. Bulk data
+  \caption{NCD operates over a simple request-reply protocol. Bulk data
     immediately follows formal messages, as required.}
@@
 \texttt{INET} sockets. All request types are encapsulated within a
-single request structure (see section \ref{sec:implementation}).
+single request structure (see Section \ref{sec:implementation}).
 Similarly, there is only a single reply structure to account for all
@@
 Because there is only one type of packet for a request, and another for
-a reply, (listing \ref{lst:protocol}) there is a small amount of wasted
+a reply, (Listing \ref{lst:protocol}) there is a small amount of wasted
 space on some file operations. We have done our best to reduce this
@@
 of including everything within a kernel module are discussed in section
-\ref{sec:currentstatus}.}. A typical userspace program can be sloppy
+\ref{sec:futurework}.}. A typical userspace program can be sloppy
 with cleanup because all memory is freed on termination. In contrast, a
@@
 Some of the basic features of the NCD kernel module are listed in
-listing \ref{lst:client}. NCD dynamically requests a major and
+Listing \ref{lst:client}. NCD dynamically requests a major and
 minor number from the running kernel \footnote{Statically assigned
@@
 
-In the snippet given in listing \ref{lst:server}, a \texttt{read()} file
+In the snippet given in Listing \ref{lst:server}, a \texttt{read()} file
 operation is handled. A buffer must be allocated with a size
@@
 Currently we are using \texttt{udev} to manage the
-permissions of imported devices (listing \ref{lst:udevrule}).
+permissions of imported devices (Listing \ref{lst:udevrule}).
 
@@
 For more functionality, imported character devices must have certain
-attributes such as those for the keyboard in listing \ref{lst:udevattr}.
+attributes such as those for the keyboard in Listing \ref{lst:udevattr}.
 Unfortunately setting them retroactively does not convince all programs
@@
 This set of core functionality provides a working proof-of-concept
-Network Character Devices system. This system fully supports the
+Network Character Device system. This system fully supports the
 export of a single character device from server to client over the
@@
 assumption. Thus, one of the challenges involved in implementing an NCD
-system is to remove the OS's assumption that any connected devices in
+system is to remove the OS's assumption that any connected device is
 available for local use. Modern-day plug-and-play auto-configuration
@@
 \item Add device protection and removal capabilities to server system
-\item Add advertisement protocol
+\item Add advertisement protocol and unreliable data protocol option
 \item Separate networking and memory-intensive code from client kernel module
@@
 
-The current NCD system only support a single device per server and per
+The current NCD system only supports a single device per server and per
 client. To be truly useful and to enable the construction of the
@@
 
+Currently, all NCD traffic is sent via TCP. While control messages
+should always use TCP to insure system state and stability, there are
+device types who generate data more appropriate for UDP operation (web
+cams, etc). In the future, we hope to allow the user to specify
+whether the data traffic for a given device will be delivered via TCP
+or whether it can be delivered using UDP.
+
 Since large scale NCD deployments may include thousands of exported
@@
 network access and memory-intensive processing directly in the kernel
-code. For security reasons, it is desirable to separate the direct
+code. This approach was taken in our initial implementation to control
+complexity.
+For security reasons, it is desirable to move the direct
 network communication from the kernel code to user-space code where
@@
 and other areas where a user wishes to use a single keyboard, video,
-and mouse, to control multiple machines. KVM systems range from simple KVM
+and mouse to control multiple machines. KVM systems range from simple KVM
 switches that allow the user to select which local machine is
@@
 
-Another example of the novel systems that can be built use Network
+Another example of the novel systems that can be built using Network
 Charter Devices is a distributed random number generator. Linux has a
@@
 timings, and other high entropy events) to generate random
-numbers. This, the speed at which random numbers can be generated is
+numbers. Thus, the speed at which random numbers can be generated is
 limited on any single system to the locally available
@@
 all imported by a single client running some from on RNG aggregator
-that combines the relatively slow output form the local RNGs into a
+that combines the relatively slow output from the local RNGs into a
 single high speed RNG output. We can effectively pool the entropy
@@
 capabilities to be exported from where they are not needed and imported
-where they are. By separating device usage from the device physical
+where they are. By separating device usage from the device's physical
 location and host, the NCD system provides a generalized, transparent,
